Make any stupid plot you want

### 2020-09-10 draft/intro.tex
munzner encoders

\documentclass[letterpaper,onecolumn,titlepage]{Ythesis}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{array,multirow}
\usepackage{subcaption}
\usepackage{subfiles}
\usepackage{url}
\usepackage{amsmath}
\usepackage{float}
\usepackage{graphicx}
\graphicspath{{figures}{.figures}}
\usepackage[backend=bibtex, style=numeric-comp]{biblatex}
\bibliography{glasslab_viz}

@@

 What even is encoding?
 \begin{figure}
    \includegraphics{figures/intro/retinal_variables}
    \label{fig:retinal_variable}
    \caption{This codification of Bertin's retinal variables \cite{bertinSemiologyGraphicsDiagrams2011} that lays out the geometry (columns) versus the aesthetics (rows) with a recommendation based on data type (qualitative versus quantative) comes from Krygier and Wood's Making Maps \cite{krygierMakingMapsVisual2005}}
\end{figure}

When talking about encodings, we are referering to Bertin's codification of the properties of the graphic system \cite{bertinSemiologyGraphicsDiagrams2011}.

\cite{bertinSemiologyGraphicsDiagrams2011}

 By encoding, we mean

\begin{figure}
    \includegraphics{figures/intro/munzner_datatypes}
    \label{fig:munzner_datatypes}
    \caption{Keys are unique lookup values used to find individual observations in the dataset. Keys are positional references, and can be coordinates on a map or unique values such as a primary key in a database or a (time, latitude, longitude) index in a data cube. Image modified from a diagram from Munzner's website\cite{munznerChDataAbstraction}}
\end{figure}
Munzner's key/value semantics\cite{munznerWhatDataAbstraction2014}
provide a way to identify variables that in effect act as metadata for other variables, such as how when we are interested in the temperature at a time the temperature is the lookup value. But

 that did not generalize well for complex heterogenous datasets. In this proposal, we refine those ideas by using topology to formally describe the connectivity between the measurements.

>>>>

\documentclass[letterpaper,onecolumn,titlepage]{Ythesis}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{array,multirow}
\usepackage{subcaption}
\usepackage{subfiles}
\usepackage{url}
\usepackage{amsmath}
\usepackage{float}
\usepackage{graphicx}
\usepackage[backend=bibtex, style=numeric-comp]{biblatex}
\bibliography{glasslab_viz}

@@

 What even is data?
 Remake this graph notated w/ graph semantics:
 \begin{figure}
    \includegraphics{figures/intro/data_formatting}
    \caption{Modified from Munzner's Visualizarion Analysis and Design}
 \end{figure}

 What even is encoding?
 \begin{figure}
    \includegraphics{figures/intro/retinal_variables}
    \label{fig:retinal_variable}
    \caption{This organization of Bertin's retinal variables \cite{bertinSemiologyGraphicsDiagrams2011} that lays out the geometry(marks) and corresponding aesthetic variations (channels) with a recommendation based on data type (qualitative versus quantative) comes from Krygier and Wood's Making Maps \cite{krygierMakingMapsVisual2005}}
\end{figure}

When talking about encodings, we are referring to Bertin's codification of the properties of the graphic system \cite{bertinIIPropertiesGraphic2011} seen in fig~\ref{fig:retinal_variable} as visual variables. In this system, there are the marks, which are the point, line, and area geometric primatives on the visual plane that each represent a single instance of the data being visualized \cite{bertinIIPropertiesGraphic2011,munznerMarksChannels2014}. These marks can be varied along different visual channels  size, value, texture, color, orientation, shape, and position \cite{bertinIIPropertiesGraphic2011, munznerMarksChannels2014}. As shown in fig~\ref{fig:retinal_variable}, the marks encode topological connectivity such that they have the following graph semantics:

\begin{description} 
    \item[Points] - 0d, no edges
    \item[Lines] - 1d, every vertex has <=2 edges (1d simplex)
    \item[Area] - 2d,  every vertex has <=4 edges (2d simplex)
\end{description}

The visual channels in fig~\ref{fig:retinal_variable} are recommended based on the measurement type, quantative versus qualitativem  and Munzner generalizes the quantative channels as magnitude channels for ordered attributes (measurements drawn from an ordered set of possible values) and the qualitative channels as identity channels for categorical attributes (random set of possible values).

\begin{figure}
    \includegraphics{figures/intro/munzner_datatypes}
    \label{fig:munzner_datatypes}
    \caption{Keys are unique lookup values used to find individual observations in the dataset. Keys are positional references, and can be coordinates on a map or unique values such as a primary key in a database or a (time, latitude, longitude) index in a data cube. Image modified from a diagram from Munzner's website\cite{munznerChDataAbstraction}}
\end{figure}
Munzner's key/value semantics\cite{munznerWhatDataAbstraction2014}
provide a way to identify variables that in effect act as metadata for other variables, such as how when we are interested in the temperature at a time the temperature is the lookup value. But that did not generalize well for complex heterogenous datasets. In this proposal, we refine those ideas by using topology to formally describe the connectivity between the measurements.
